---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# deltamapr

<!-- badges: start -->
[![R build status](https://github.com/InteragencyEcologicalProgram/deltamapr/workflows/R-CMD-check/badge.svg)](https://github.com/InteragencyEcologicalProgram/deltamapr/actions)
<!-- badges: end -->

The goal of deltamapr is to provide spatial data for the Bay-Delta.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("InteragencyEcologicalProgram/deltamapr")
```

## Usage

This package contains data objects stored in [`sf`](https://r-spatial.github.io/sf/) format, which can be accessed with:
```{r}
require(sf) # To ensure `sf` objects print correctly
deltamapr::WW_Delta
```

Or, after loading the package, like this:

```{r}
require(deltamapr)
WW_Delta
```


## Included data sets

```{r, message=FALSE}
require(ggplot2)
require(sf)
```

### Delta waterways

```{r}
ggplot(WW_Delta)+
  geom_sf()+
  theme_bw()
```

### Full watershed waterways

```{r}
ggplot(WW_Watershed)+
  geom_sf()+
  theme_bw()
```


### Delta regions

```{r}
ggplot(R_EDSM_Mahardja)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### [California aquatic resources inventory](https://www.sfei.org/data/california-aquatic-resource-inventory-cari-version-03-gis-data#sthash.zuIpGDaQ.dpbs)

#### CARI polygons

```r
ggplot() +
  geom_sf(data = SFE_CARI_polygons, aes(fill = leglabel1))
```

#### CARI streams

```r
ggplot() +
  geom_sf(data = SFE_CARI_streams)
```
