---
title: deltamapr
output: 
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![R build status](https://github.com/InteragencyEcologicalProgram/deltamapr/workflows/R-CMD-check/badge.svg)](https://github.com/InteragencyEcologicalProgram/deltamapr/actions)
<!-- badges: end -->

The goal of deltamapr is to provide spatial data for the Bay-Delta.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("InteragencyEcologicalProgram/deltamapr")
```

## Usage

This package contains data objects stored in [`sf`](https://r-spatial.github.io/sf/) format, which can be accessed with:
```{r}
require(sf) # To ensure `sf` objects print correctly
deltamapr::WW_Delta
```

Or, after loading the package, like this:

```{r}
require(deltamapr)
WW_Delta
```

## Data types

Three spatial data types are included in this package

1) **Waterways**: Shapefiles of water coverage. These data files are preceded by the prefix *WW*.
2) **Regions**: Shapefiles with sets of regions, subregions, or strata. These data files are preceded by the prefix *R*.
3) **Habitat**: Shapefiles with habitat data. These data files are preceded by the prefix *H*.

# Waterways

```{r, message=FALSE}
require(ggplot2)
require(sf)
```

## Delta waterways

Produced by the California Department of Fish and Wildlife

```{r WWDelta}
ggplot(WW_Delta)+
  geom_sf()+
  theme_bw()
```

## DBW waterways

Produced by the California State Parks Division of Boating and Waterways

```{r WWDBW}
ggplot(WW_DBW)+
  geom_sf()+
  theme_bw()
```

## Full watershed waterways

This shapefile includes data from CH2M Hill, and Wetland and Water Resources, CDFG Ecosystem Restoration Program as a part of ERP Stage 2 conservation strategy, and Dan Gillenwater.

```{r WWWatershed}
ggplot(WW_Watershed)+
  geom_sf()+
  theme_bw()
```

# Regions

All regions are derived from the united States Fish and Wildlife Service [Enhanced Delta Smelt Monitoring Program](https://www.fws.gov/lodi/juvenile_fish_monitoring_program/jfmp_index.htm).

## Regions

### EDSM 2016-17 Phase 1 Regions (Equivalent to 2017 Phases 2&3 Regions.)

```{r REDSMRegions1617P1}
ggplot(R_EDSM_Regions_1617P1)+
  geom_sf(aes(fill=Region))+
  theme_bw()
```

### EDSM 2017-18 Phase 1 Regions

```{r REDSMRegions1718P1}
ggplot(R_EDSM_Regions_1718P1)+
  geom_sf(aes(fill=Region))+
  theme_bw()
```

### EDSM 2018 Phases 2-3 Regions (Equivalent to 2019 Phase 2 and 2020 Phase 2-3 Regions)

```{r REDSMRegions18P23}
ggplot(R_EDSM_Regions_18P23)+
  geom_sf(aes(fill=Region))+
  theme_bw()+
  theme()
```

### EDSM 2018-19 Phase 1 Regions (Equivalent to 2019-2020 Phase 1 Regions)

```{r REDSMRegions1819P1}
ggplot(R_EDSM_Regions_1819P1)+
  geom_sf(aes(fill=Region))+
  theme_bw()
```

### EDSM 2019 Phase 3 Regions

```{r REDSMRegions19P3}
ggplot(R_EDSM_Regions_19P3)+
  geom_sf(aes(fill=Region))+
  theme_bw()
```

## Strata

### EDSM 2016-17 Phase 1 Strata

```{r REDSMStrata1617P1}
ggplot(R_EDSM_Strata_1617P1)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()+
  theme()
```

### EDSM 2017 Phase 2 Strata

```{r REDSMStrata17P2}
ggplot(R_EDSM_Strata_17P2)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```

### EDSM 2017 Phase 3 Strata

```{r REDSMStrata17P3}
ggplot(R_EDSM_Strata_17P3)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```

### EDSM 2017-18 Phase 1 Strata

```{r REDSMStrata1718P1}
ggplot(R_EDSM_Strata_1718P1)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```

### EDSM 2018 Phases 2-3 Strata (Equivalent to 2019 Phase 2 and 2020 Phase 2-3 Strata)

```{r REDSMStrata18P23}
ggplot(R_EDSM_Strata_18P23)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```

### EDSM 2018-19 Phase 1 Strata (Equivalent to 2019-2020 Phase 1 Strata)

```{r REDSMStrata1819P1}
ggplot(R_EDSM_Strata_1819P1)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```

### EDSM 2019 Phase 3 Strata

```{r REDSMStrata19P3}
ggplot(R_EDSM_Strata_19P3)+
  geom_sf(aes(fill=Stratum))+
  theme_bw()
```


## Subregions

### EDSM 2016-17 Phase 1 Subregions (Equivalent to 2017 Phase 3 Subregions)

```{r REDSMSubregions1617P1}
ggplot(R_EDSM_Subregions_1617P1)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### EDSM 2017 Phase 2 Subregions

```{r REDSMSubregions17P2}
ggplot(R_EDSM_Subregions_17P2)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### EDSM 2017-18 Phase 1 Subregions

```{r REDSMSubregions1718P1}
ggplot(R_EDSM_Subregions_1718P1)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### EDSM 2018 Phases 2-3 Subregions (Equivalent to 2019 Phase 2 and 2020 Phase 2-3 Subregions)

```{r REDSMSubregions18P23}
ggplot(R_EDSM_Subregions_18P23)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### EDSM 2018-19 Phase 1 Subregions (Equivalent to 2019-2020 Phase 1 Subregions)

```{r REDSMSubregions1819P1}
ggplot(R_EDSM_Subregions_1819P1)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### EDSM 2019 Phase 3 Subregions

```{r REDSMSubregions19P3}
ggplot(R_EDSM_Subregions_19P3)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

### Brian Mahardja's edited EDSM subregions

```{r REDSMSubregionsMahardja}
ggplot(R_EDSM_Subregions_Mahardja)+
  geom_sf(aes(fill=SubRegion))+
  theme_bw()+
  theme(legend.position="none")
```

# Habitat

## California aquatic resources inventory

San Francisco Estuary Institute (SFEI). 2017. "California Aquatic Resource Inventory (CARI) version 0.3." Accessed Aug 26, 2020. http://www.sfei.org/data/california-aquatic-resource-inventory-cari-version-03-gis-data

### CARI wetlands

```{r HCARIwetlands, fig.height=12}
ggplot(H_CARI_wetlands) +
  geom_sf(aes(fill = leglabellevel1))+
  scale_fill_discrete(name="Wetland type", guide=guide_legend(ncol=1, title.position = "top", label.theme=element_text(size=8)))+
  theme_bw()+
  theme(legend.position = "bottom")
```

### CARI streams

```{r HCARIstreams}
ggplot(H_CARI_streams) +
  geom_sf()+
  theme_bw()
```
